Open Data, Open Source and Open Standards in chemistry: \\ The Blue Obelisk five years on

### 2011-05-18 paper.tex
Expanded JNI-InChI section

The IUPAC InChI identifier is a non-proprietary and unique identifier
for chemical substances designed to enable linking of diverse data
compilations. Although its development predates the Blue Obelisk,
software such as Open Babel has included InChI support since 2005,
and support for InChI in Indigo is due in 2011.

Since the official InChI implementation is in C, it is difficult to
access from the other widely used language for cheminformatics
toolkits, Java. The Blue Obelisk project JNI-InChI~\cite{WebJNIInChI}
was established in 2006 to
solve this problem by using the Java Native Interface to link the
InChI binary to Java. In this way, it promotes the wider adoption of
this standard identifier by the chemistry community.

>>>>

The IUPAC InChI identifier is a non-proprietary and unique identifier
for chemical substances designed to enable linking of diverse data
compilations. Prior to the development of the InChI identifier chemical
information systems and databases used a wide variety of (generally
proprietary) identifiers, greatly limiting their interoperability.
Although its development predates the Blue Obelisk, software such as Open
Babel has included InChI support since 2005, and support for InChI in
Indigo is due in 2011.

Since the official InChI implementation is in C, it is difficult to
access from the other widely used language for cheminformatics
toolkits, Java. Early attempts to generate InChI identifiers from
within Java involved programatically launching the InChI executable
and capturing the output, an approach that was found to be fairly
unreliable and broke the 'write once, run anywhere' philosophy
of Java.  The Blue Obelisk project JNI-InChI~\cite{WebJNIInChI}
was established in 2006 to solve this problem by using the Java Native
Interface framework to provide transparent access to the InChI
library from within Java, and other Java Virtual Machine (JVM) based
languages, supporting the wider adoption of
this standard identifier by the chemistry community.

The Java Native Interface framework provides a mechanism for code
running inside the JVM to place calls to libraries written in languages
such as C, C++ and Fortran, and compiled into native, machine specific,
code. JNI-InChI provides a thin C wrapper, with corresponding Java code,
around the IUPAC InChI library, exposing the InChI library's functionality
to the JVM.  To overcome the need to have the correct InChI library pre-installed
on a system, JNI-InChI comes with a variety of precompiled native binaries and
automatically extracts and deploys the correct one for the detected operating
system and architecture. The JNI-InChI library comes with native binaries
supporting a range of operating systems and architectures; the current version
has binaries for 32- and 64-bit Windows, Linux and Solaris, 64-bit FreeBSD
and 64-bit Intel-based Mac OS X - a number of which are not supported by the
original IUPAC distribution of InChI.

The JNI-InChI project has matured to support the full range of
functionality of the InChI C library: structure-to-InChI, InChI-to-structure,
AuxInfo-to-structure, InChIKey generation, and InChI and InChIKey validation.
JNI-InChI provides the InChI functionality for a number of Open Source projects,
including the Chemistry Development Kit, Bioclipse and CMLXOM/JUMBO, and is
also used by commercial applications and internally in a number of companies.
Through its widespread use and Open Source development model, a number of
issues in earlier versions of the software have been identified and resolved,
and JNI-InChI now offers a robust, and defacto standard tool for working with
InChIs in the JVM.
